---
title: "Divvytrips_Case_Study"
author: "Brian Allen"
date: "2024-05-16"
output:
  pdf_document: default
  html_document: default
---
# Detailed Analysis

Cyclistic is looking for marketing strategies to convert casual riders into annual members.

## Questions to answer about the data

1. How do members and casuals differ from each other?
2. Is there a 'why' in the data for casual riders to convert?
3. How can marketing be used to highlight the 'why' and influence casuals?

### Loading Packages
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(readr)
library(ggplot2)
library(janitor)
library(lubridate)
library(vroom)
library(geosphere)
options(scipen = 9)

#Read the divvy_data csv from our cleaning
divvy_data = read.csv("C:/Users/barry/~coding/R Programming/googleclass/divvytrips/divvy_data.csv")
divvy_data_time = read.csv("C:/Users/barry/~coding/R Programming/googleclass/divvytrips/divvy_data_time.csv")
```

My hypothesis is that the classic bikes and electric bikes may have differences 
when it comes to comparing members and casuals. 

I also want to add a column that calculates total ride time. I think that there
may be a difference with length of ride.

I also want to calculate the most frequent routes and specifically look at
routes that are frequent between Mon-Fri as well as the weekend Sat-Sun.

First lets test the mean ride times and top 5 starting and ending stations by member type. 

```{r}

divvy_data %>% 
  group_by(member_casual,rideable_type) %>% 
  summarise(mean_ride = mean(total_time))

```


```{r}
divvy_data %>% 
  filter(member_casual == 'casual' & start_station_id == "13022" & rideable_type == 'classic_bike') %>%
  summarise(start_station = n())
```

Separating the data by bike type to simplify comparison and reduce computation
time on my personal computer. 

```{r Classic_bikes}
classic_data = divvy_data %>% 
  filter(rideable_type == 'classic_bike')

head(classic_data)
```

```{r Electric_bikes}
electric_data = divvy_data %>% 
  filter(rideable_type == 'electric_bike')

head(electric_data)
```

```{r docked_bikes}
docked_data = divvy_data %>% 
  filter(rideable_type == 'docked_bike')

head(docked_data)
```
Find most frequent start stations. 


```{r Classic_top_casual_start}
classic_data %>% 
  filter(member_casual == 'casual') %>%
  group_by(start_station_id,start_station_name) %>% 
  summarise(start_station = n()) %>% 
  arrange(desc(start_station)) %>% 
  head(n=10)
```

```{r Classic_top_member_start}
classic_data %>% 
  filter(member_casual == 'member') %>%
  group_by(start_station_id,start_station_name) %>% 
  summarise(start_station = n()) %>% 
  arrange(desc(start_station)) %>% 
  head(n=10)
```
Top end stations. 

```{r Classic_top_casual_end}
classic_data %>% 
  filter(member_casual == 'casual') %>%
  group_by(end_station_id,end_station_name) %>% 
  summarise(end_station = n()) %>% 
  arrange(desc(end_station)) %>% 
  head(n=10)
```


```{r Classic_top_member_end}
classic_data %>% 
  filter(member_casual == 'member') %>%
  group_by(end_station_id,end_station_name) %>% 
  summarise(end_station = n()) %>% 
  arrange(desc(end_station)) %>% 
  head(n=10)
```
electric

```{r electric_top_member_start}
electric_data %>% 
  filter(member_casual == 'member') %>%
  group_by(start_station_id,start_station_name) %>% 
  summarise(start_station = n()) %>% 
  arrange(desc(start_station)) %>% 
  head(n=10)
```

```{r electric_top_member_end}
electric_data %>% 
  filter(member_casual == 'member') %>%
  group_by(end_station_id,end_station_name) %>% 
  summarise(end_station = n()) %>% 
  arrange(desc(end_station)) %>% 
  head(n=10)
```


With the start and end stations being similar for the top options.
I am curious how many trips start and stop at the same station and if there is
a difference between members and casuals. 
```{r}
electric_data %>% 
  filter(start_station_id == end_station_id) %>%
  group_by(start_station_name,end_station_name, member_casual) %>% 
  summarise(trips = n()) %>% 
  arrange(desc(trips)) %>% 
  head(n=10)
```

how many trips start and stop at the same station?
```{r}
divvy_data %>% 
  filter(start_station_id == end_station_id) %>% 
  group_by(start_station_name,end_station_name, member_casual, rideable_type) %>% 
  summarise(trips = n()) %>% 
  arrange(desc(trips)) %>% 
  head(n=10)
```
How many trips start and stop at different locations?
```{r}
divvy_data %>% 
  filter(start_station_id != end_station_id) %>% 
  group_by(start_station_name,end_station_name, member_casual, rideable_type) %>% 
  summarise(trips = n()) %>% 
  arrange(desc(trips)) %>% 
  head(n=10)
```
What is interesting is that the largest frequency of start/end stations
being the same are by Casual riders while the highest frequency of different 
start/end stations are by member riders. 

This implies that casual riders ride around and return the bike to the same
station they took the bike from. While member riders are point A to point B.

Lets see if there is a difference in ride time. Start with top then average

```{r}
divvy_data %>% 
  filter(start_station_id == end_station_id) %>% 
  select(start_station_name,end_station_name, member_casual,
           rideable_type, total_time) %>% 
  arrange(desc(total_time)) %>% 
  head(n=10)
```

```{r}
divvy_data %>% 
  filter(start_station_id != end_station_id) %>% 
  select(start_station_name,end_station_name, member_casual,
           rideable_type, total_time) %>% 
  arrange(desc(total_time)) %>% 
  head(n=10)
```
Lets see the averages
```{r}
divvy_data %>% 
  filter(start_station_id == end_station_id) %>% 
  group_by(member_casual) %>% 
  summarise(average_ride_time = mean(total_time))
```
```{r}
divvy_data %>% 
  filter(start_station_id != end_station_id) %>% 
  group_by(member_casual) %>% 
  summarise(average_ride_time = mean(total_time))
```



```{r}
ggplot(data=divvy_data) +
  geom_bar(mapping = aes(x= factor(month_abb,levels=month.abb),fill=member_casual))+
  xlab("Months")+
  facet_wrap(~rideable_type, ncol = 1) +
  theme(axis.text.x = element_text(angle = 45))
```
At this point I am thinking that the docked bikes are irrelevant. 

Compare ride mix weekends vs weekdays.
```{r}
start_date = as.POSIXct(min(divvy_data$started_at))
end_date = as.POSIXct(max(divvy_data$ended_at))

#start_date = "2023-05-01 00:00:33"
#end_date = "2024-05-01 06:03:21"

num_weekdays = sum(!weekdays(seq(from=start_date, to=end_date, by="days")) %in% c("Saturday","Sunday"))
num_weekends = sum(weekdays(seq(from=start_date, to=end_date, by="days")) %in% c("Saturday","Sunday"))

divvy_data %>% 
  summarize(cust = "casual", total_weekend_trips = sum(divvy_data$day_week %in% c("Saturday","Sunday")
                                      & divvy_data$member_casual == "casual"),
            total_weekdays_trips = sum(!divvy_data$day_week %in% c("Saturday","Sunday")
                                       & divvy_data$member_casual == "casual"),
            avg_weekend_trips = sum(divvy_data$day_week %in% c("Saturday","Sunday")
                                     & divvy_data$member_casual == "casual")/num_weekends,
            avg_weekday_trips = sum(!divvy_data$day_week %in% c("Saturday","Sunday")
                                     & divvy_data$member_casual == "casual")/num_weekdays,
            sat_sun = num_weekends, mon_fri = num_weekdays)

divvy_data %>% 
  summarize(cust = "member", total_weekend_trips = sum(divvy_data$day_week %in% c("Saturday","Sunday")
                                      & divvy_data$member_casual == "member"),
            total_weekdays_trips = sum(!divvy_data$day_week %in% c("Saturday","Sunday")
                                       & divvy_data$member_casual == "member"),
            avg_weekend_trips = sum(divvy_data$day_week %in% c("Saturday","Sunday")
                                     & divvy_data$member_casual == "member")/num_weekends,
            avg_weekday_trips = sum(!divvy_data$day_week %in% c("Saturday","Sunday")
                                     & divvy_data$member_casual == "member")/num_weekdays,
            sat_sun = num_weekends, mon_fri = num_weekdays)
```


```{r}
divvy_data %>% 
  filter(day_week %in% c("Saturday","Sunday")) %>% 
  ggplot() +
  geom_bar(mapping = aes(x= factor(month_abb,levels=month.abb),fill=rideable_type))+
  xlab("Months")+
  facet_wrap(~member_casual, ncol =1) +
  theme(axis.text.x = element_text(angle = 45))
```
Weekend distribution
```{r}
divvy_data %>% 
  filter(day_week %in% c("Saturday","Sunday")) %>% 
  ggplot() +
  geom_bar(mapping = aes(x= factor(month_abb,levels=month.abb),fill=member_casual))+
  xlab("Months")
```
Weekend or not weekend
```{r}
options(scipen = 9)

divvy_data %>% 
  ggplot() +
  geom_bar(mapping = aes(x= factor(month_abb,levels=month.abb),fill=member_casual))+
  xlab("Months") +
  facet_wrap(~day_week %in% c("Saturday","Sunday"), ncol=1)
```

```{r}
divvy_data %>% 
  group_by(member_casual) %>% 
  summarize(avg_distance = mean(distance),avg_time = mean(total_time), 
            zero_distance = sum(distance == 0), has_distance = sum(distance != 0),
            min_distance = min(distance), max_distance = max(distance))
  
```

```{r}
divvy_data %>% 
  filter(distance > 0) %>% 
  group_by(member_casual) %>% 
  summarize(avg_distance = mean(distance),avg_time = mean(total_time), 
            zero_distance = sum(distance == 0), has_distance = sum(distance != 0),
            min_distance = min(distance), max_distance = max(distance))
  
```

```{r}
divvy_data %>% 
  filter(distance == 0) %>% 
  group_by(member_casual) %>% 
  summarize(avg_distance = mean(distance),avg_time = mean(total_time), 
            zero_distance = sum(distance == 0), has_distance = sum(distance != 0),
            min_distance = min(distance), max_distance = max(distance))
  
```

```{r}
options(scipen = 9)
histvecbrk = seq(0,5000, by=500)

divvy_data %>% 
  filter(distance < 5000 & distance > 0 & member_casual == "casual") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 500, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Casual Riders",
    subtitle = "Less than 5,000m excluding 0m",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 500) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )

```
```{r}
divvy_data %>% 
  filter(distance < 5000 & distance > 0 & member_casual == "member") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 500, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Member Riders",
    subtitle = "Less than 5,000m excluding 0m",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 500) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
  
```
Looking at the distribution of time. 

```{r}
divvy_data %>% 
  filter(distance < 5000 & total_time < 3601 & member_casual == "casual") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 500, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Casual Riders",
    subtitle = "Less than 5,000m and less than 3hrs",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 500) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
```

```{r}
divvy_data %>% 
  filter(distance < 5000 & total_time < 3601 & member_casual == "member") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 500, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Member Riders",
    subtitle = "Less than 5,000m and less than 3hrs",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 500) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
```
```{r}
divvy_data %>% 
  group_by(member_casual) %>% 
  summarize(rides = sum(distance < 2500 & total_time < 3601))
```

```{r}
histvecbrk2 = seq(0,2500, by=250)

divvy_data %>% 
  filter(distance < 2500 & total_time < 3601 & member_casual == "member") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 250, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Member Riders",
    subtitle = "Less than 2,500m and less than 3hrs",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 250) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk2) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
```



```{r}
histvecbrk2 = seq(0,2500, by=250)

divvy_data %>% 
  filter(distance < 2500 & total_time < 3601 & member_casual == "casual") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 250, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Casual Riders",
    subtitle = "Less than 2,500m and less than 3hrs",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 250) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk2) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
```

```{r}
divvy_data %>% 
  group_by(member_casual) %>% 
  summarize(rides = sum(distance < 2500 & total_time < 3601 & day_week == "Saturday" | distance < 2500 & total_time < 3601 & day_week == "Sunday"))
```

```{r}
divvy_data %>% 
  filter(member_casual == "casual" & distance < 2500 & total_time < 3601 & day_week == "Saturday" | 
           member_casual == "casual" & distance < 2500 & total_time < 3601 & day_week == "Sunday") %>% 
  ggplot(aes(x=distance, y = after_stat(count / sum(count)))) + 
  geom_histogram(binwidth = 250, color = "#000000", fill = "#0099F8")+ 
  labs(
    title = "Distribution of Weekend Casual Riders",
    subtitle = "Less than 2,500m and less than 3hrs",
    caption = "Source: divvytrips data",
    x = "Distance Meters",
    y = "Distribution %"
  ) +
  stat_bin(aes(label = scales::percent(stat(round((count / sum(count)),2)))),geom = "text", vjust = 1.1, binwidth = 250) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = histvecbrk2) +
  theme_classic() +
  theme(
    plot.title = element_text(color = "#0099F8", size = 14, face = "bold"),
    plot.subtitle = element_text(size = 8, face = "bold"),
    plot.caption = element_text(face = "italic"),
    axis.text.x = element_text(angle = 45)
    )
```


```{r}
options(scipen = 9)

divvy_data %>% 
  filter(distance < 5000 & distance > 0) %>% 
  ggplot(aes(x=distance)) +
  geom_histogram()+
  xlim(0,5000) + 
  ylim(0,225000) +
  facet_wrap(~member_casual)
```




```{r}
options(scipen = 9)

divvy_data %>% 
  filter(distance == 0 & total_time < 3600 ) %>% 
  ggplot(aes(x=total_time / 60)) +
  geom_histogram(binwidth = 5, fill = "#4fb9af", color = "black") + 
  facet_wrap(~member_casual)
```

Lets use electric bikes as an example of a marketing opportunity.
Using the costing of electric bikes, we can compare the potential savings
that a casual rider could receive. Assumption is that there are casual riders
using bikes at a similar frequency to members. Potential commuters. 



```{r}
time0 = sum(divvy_data_time$total_time >= 120 & divvy_data_time$total_time < 299)

time3 = sum(divvy_data_time$total_time > 299 & divvy_data_time$total_time < 479)

time6 = sum(divvy_data_time$total_time > 479 & divvy_data_time$total_time < 659)

time9 = sum(divvy_data_time$total_time > 659 & divvy_data_time$total_time < 839)

time12 = sum(divvy_data_time$total_time > 839 & divvy_data_time$total_time < 1019)

time15 = sum(divvy_data_time$total_time > 1019 & divvy_data_time$total_time < 1199)

time18 = sum(divvy_data_time$total_time > 1199 & divvy_data_time$total_time < 1379)

time21 = sum(divvy_data_time$total_time > 1379 & divvy_data_time$total_time < 1559)

time24 = sum(divvy_data_time$total_time > 1559 & divvy_data_time$total_time < 1739)

time27 = sum(divvy_data_time$total_time > 1739 & divvy_data_time$total_time < 1919)

#casual

time0_casual = sum(divvy_data_time$member_casual == "casual" &
                     divvy_data_time$total_time >= 120 & divvy_data_time$total_time < 299)

time3_casual = sum(divvy_data_time$member_casual == "casual" &
                     divvy_data_time$total_time > 299 & divvy_data_time$total_time < 479)

time6_casual = sum(divvy_data_time$member_casual == "casual" &
                     divvy_data_time$total_time > 479 & divvy_data_time$total_time < 659)

time9_casual = sum(divvy_data_time$member_casual == "casual" &
                     divvy_data_time$total_time > 659 & divvy_data_time$total_time < 839)

time12_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 839 & divvy_data_time$total_time < 1019)

time15_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 1019 & divvy_data_time$total_time < 1199)

time18_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 1199 & divvy_data_time$total_time < 1379)

time21_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 1379 & divvy_data_time$total_time < 1559)

time24_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 1559 & divvy_data_time$total_time < 1739)

time27_casual = sum(divvy_data_time$member_casual == "casual" &
                      divvy_data_time$total_time > 1739 & divvy_data_time$total_time < 1919)

#member

time0_member = sum(divvy_data_time$member_casual == "member" &
                     divvy_data_time$total_time >= 120 & divvy_data_time$total_time < 299)

time3_member = sum(divvy_data_time$member_casual == "member" &
                     divvy_data_time$total_time > 299 & divvy_data_time$total_time < 479)

time6_member = sum(divvy_data_time$member_casual == "member" &
                     divvy_data_time$total_time > 479 & divvy_data_time$total_time < 659)

time9_member = sum(divvy_data_time$member_casual == "member" &
                     divvy_data_time$total_time > 659 & divvy_data_time$total_time < 839)

time12_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 839 & divvy_data_time$total_time < 1019)

time15_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 1019 & divvy_data_time$total_time < 1199)

time18_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 1199 & divvy_data_time$total_time < 1379)

time21_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 1379 & divvy_data_time$total_time < 1559)

time24_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 1559 & divvy_data_time$total_time < 1739)

time27_member = sum(divvy_data_time$member_casual == "member" &
                      divvy_data_time$total_time > 1739 & divvy_data_time$total_time < 1919)
#cost benefit

time0_bene = ((2 *  0.44) - (2 * 0.18)) * time0_casual

time3_bene = ((5 *  0.44) - (2 * 0.18)) * time3_casual

time6_bene = ((8 *  0.44) - (2 * 0.18)) * time6_casual

time9_bene = ((11 *  0.44) - (2 * 0.18)) * time9_casual

time12_bene = ((14 *  0.44) - (2 * 0.18)) * time12_casual

time15_bene = ((17 *  0.44) - (2 * 0.18)) * time15_casual

time18_bene = ((20 *  0.44) - (2 * 0.18)) * time18_casual

time21_bene = ((23 *  0.44) - (2 * 0.18)) * time21_casual

time24_bene = ((26 *  0.44) - (2 * 0.18)) * time24_casual

time27_bene = ((29 *  0.44) - (2 * 0.18)) * time27_casual

time = c(2,5,8,11,14,17,20,23,26,29)
tripscount = c(time0,time3,time6,time9,time12,time15,time18,time21,time24,time27)

casualtrips = c(time0_casual,time3_casual,time6_casual,time9_casual,time12_casual,
                time15_casual,time18_casual,time21_casual,time24_casual,time27_casual)

membertrips = c(time0_member,time3_member,time6_member,time9_member,time12_member,
                time15_member,time18_member,time21_member,time24_member,time27_member)

benefit = c(time0_bene,time3_bene,time6_bene,time9_bene,time12_bene,
                time15_bene,time18_bene,time21_bene,time24_bene,time27_bene)

timeblock_chart = data.frame(time,tripscount,casualtrips,membertrips,benefit)

```


Trip counts
```{r}
options(scipen = 9)
ggplot(timeblock_chart, aes(x = factor(time), y = tripscount)) +
  geom_bar(stat = "identity", fill = "#0099F8") +
  labs(title = "Trips Count Over Time",
       x = "Time",
       y = "Trips Count") +
  theme_minimal()
```

Casual Trip counts
```{r}
options(scipen = 9)
ggplot(timeblock_chart, aes(x = factor(time), y = casualtrips)) +
  geom_bar(stat = "identity", fill = "#0099F8") +
  labs(title = "Casual Trips Over Time",
       x = "Time",
       y = "Trips Count") +
  theme_minimal()
```
Member Trip counts
```{r}
options(scipen = 9)
ggplot(timeblock_chart, aes(x = factor(time), y = membertrips)) +
  geom_bar(stat = "identity", fill = "#0099F8") +
  labs(title = "Member Trips Over Time",
       x = "Time",
       y = "Trips Count") +
  theme_minimal()
```
Casual to member benefit assuming cummuting 5 days a week. two trips.
```{r}
options(scipen = 9)
ggplot(timeblock_chart, aes(x = factor(time), y = benefit)) +
  geom_bar(stat = "identity", fill = "#0099F8") +
  labs(title = "Benefit by Commute Time",
       x = "Time",
       y = "Benefit $") +
  theme_minimal()
```